---
title: "ILE Dashboard"
author: "Parth Jariwala"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    navbar:
      - icon: "fa-github"
        href: "https://github.com/19jariwalap/final_DATA555"
        align: "right"
      - icon: "fa-envelope"
        href: "mailto:parth.jariwala@emory.edu"
        align: "right"
        title: "Email me with any questions"
---

```{r setup, include=FALSE}
# Loading libraries
library(flexdashboard)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(tidyverse)
library(gtsummary)
library(gt)
library(knitr)
library(webshot2)
library(here)
library(plotly)
library(bkmr)
library(htmlwidgets)
```

**PFAS Boxplots**
============
Row
-----------------------------------------------------------------------

### **Overview & Key Takeaway**

This interactive chart shows levels of 14 per- and polyfluoroalkyl substances (PFAS), also known as “forever chemicals”. In my study, PFOS, PFOA, PFNA, and PFHxS (types of PFAS) had the highest concentrations in the participants. Using the lasso tool, you can select and snapshot the less abundant chemicals and switching to a log scale spreads the data out so both large and small values are easier to see.

## **PFAS Boxplots**

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Reading in data
CLU0106_PFAS_clean_half_min_new <- readRDS(here::here("data", "CLU0106_PFAS_clean_half_min_new.rds"))

PFAS_long <- CLU0106_PFAS_clean_half_min_new %>% 
  pivot_longer(cols = c("PFHxS", "PFNA", "PFOA", "PFOS", "N-MeFOSAA", 
                        "PFBS", "PFHps", "PFDA", "PFDoA", 
                        "PFHpA", "PFHxA", "PFTrDA", "PFUnA", "PFPeAS"), 
               names_to = "PFAS", values_to = "Concentration")

PFAS_long_new <- PFAS_long %>% select(sample_id, PFAS, Concentration)

# Making boxplot

plotly1 <- PFAS_long_new %>%
  plot_ly(
    x         = ~PFAS,
    y         = ~Concentration,
    type      = "box",
    boxpoints = FALSE,
    color     = ~PFAS,                                   
    colors    = viridis::viridis(n = length(unique(PFAS_long_new$PFAS)))  
  ) %>%
  layout(
    width  = 1150,
    height = 600,
    title  = list(text = "Figure 1. Distribution of PFAS Concentrations", font = list(size = 14)),
    xaxis  = list(title = "PFAS Type"),
    yaxis  = list(title = "Concentration (ng/mL)", rangeslider = list(visible = TRUE)),
    updatemenus = list(
      list(
        direction   = "down",
        showactive  = TRUE,
        x           = 0.1,
        xanchor     = "left",
        y           = 1.2,
        yanchor     = "top",
        buttons     = list(
          list(
            method = "relayout",
            args   = list(list(yaxis = list(type = "linear"))),
            label  = "Linear Scale"
          ),
          list(
            method = "relayout",
            args   = list(list(yaxis = list(type = "log"))),
            label  = "Log Scale"
          )
        )
      )
    )
  ) %>%
  config(
    displayModeBar      = TRUE,
    modeBarButtonsToAdd = c("lasso2d", "select2d")
  )


plotly1

```


**BKMR Results 1** {.tabset}
============
Row
-----------------------------------------------------------------------
### **Overview & Key Takeaway**
These four interactive charts show how cholesterol, systolic blood pressure, hemoglobin-A1C, and glucose change at different levels of per- and polyfluoroalkyl substances (PFAS), also known as “forever chemicals.” Whenever a line crosses a dashed threshold, the change is too uncertain to be meaningful.

In the cholesterol chart, hovering suggests a rise of about four points once PFAS exposure reaches the top 25%, but this increase isn’t statistically reliable. The systolic blood pressure chart likewise shows no clear or dependable trend with PFAS exposure.

Row
-----------------------------------------------------------------------
### **Cholesterol**
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Reading in data for BKMR results
PFAS_outcome_merged <- readRDS(here::here("data", "PFAS_outcome_merged.rds"))
bkmr_choles_model <- readRDS(here::here("data", "bkmr_choles_model.rds"))

# Generate overall risk summary for cholesterol
risks.overall_choles <- OverallRiskSummaries(fit = bkmr_choles_model, qs = seq(0.25, 0.75, by = 0.05), q.fixed = 0.5) 

bkmr_choles_plot <- ggplot(
  data = risks.overall_choles,
  mapping = aes(
    x = quantile,
    y = est,
    ymin = est - 1.96 * sd,
    ymax = est + 1.96 * sd,
    text = paste0(
      "Mean: ", round(est, 2), "<br>",
      "95% CI: [", round(est - 1.96 * sd, 2), ", ", round(est + 1.96 * sd, 2), "]"
    )
  )
) +
  geom_hline(yintercept = 0, lty = 2, col = "brown") +
  geom_pointrange() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 8)) +
  xlab("Quantiles of Joint Exposures") +
  ylab("Overall PFAS Mixture Effect on Total Cholesterol")

plotly2_choles <- ggplotly(bkmr_choles_plot, tooltip = "text") %>% 
  layout(
    title = list(
      text = "Figure 2. Total Effect of PFAS Mixture on Total Cholesterol Levels",
      font = list(size = 12)
    ),
    xaxis = list(title = "Quantiles of Joint Exposure"),
    yaxis = list(title = "Overall Effect (No unit)"),
    updatemenus = list(
      list(
        type = "slider",
        active = 3,
        x = -0.1, xanchor = "left", y = 1.15, yanchor = "top",
        currentvalue = list(prefix = "Opacity: "),
        steps = lapply(seq(0.2, 1, by = 0.2), function(op) {
          list(method = "restyle", args = list("marker.opacity", op), label = paste0(op * 100, "%"))
        })
      )
    )
  ) %>%
  config(
    displayModeBar = TRUE,
    modeBarButtonsToAdd = c("lasso2d", "select2d")
  )

plotly2_choles

```



### **Systolic Blood Pressure (SBP)**
```{r, echo=FALSE, warning=FALSE, message=FALSE}
bkmr_SBP_model <- readRDS(here::here("data", "bkmr_SBP_model.rds"))
risks.overall_SBP <- OverallRiskSummaries(fit = bkmr_SBP_model, qs = seq(0.25, 0.75, by = 0.05), q.fixed = 0.5)

bkmr_SBP_plot <- ggplot(risks.overall_SBP, mapping = aes(
    x = quantile,
    y = est,
    ymin = est - 1.96 * sd,
    ymax = est + 1.96 * sd,
    text = paste0(
      "Mean: ", round(est, 2), "<br>",
      "95% CI: [", round(est - 1.96 * sd, 2), ", ", round(est + 1.96 * sd, 2), "]"
    )
  )
) + 
  geom_hline(yintercept = 0, lty = 2, col = "brown") +
  geom_pointrange() +    
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 8)) +
  xlab("Quantiles of Joint PFAS Exposures") +
  ylab("Overall Mixture Effect on Systolic Blood Pressure")

plotly3_SBP <- ggplotly(bkmr_SBP_plot, tooltip = "text") %>% 
  layout(
    title = list(
          text = "Figure 3. Total Effect of PFAS Mixture on Systolic Blood Pressure (SBP)",
          font = list(size = 12)
            ),
    xaxis = list(title = "Quantiles of Joint Exposure"),
    yaxis = list(title = "Overall Effect (No unit)"),
    updatemenus = list(
      list(
        type = "slider",
        active = 3,
        x = -0.1, xanchor = "left", y = 1.15, yanchor = "top",
        currentvalue = list(prefix = "Opacity: "),
        steps = lapply(seq(0.2, 1, by = 0.2), function(op) {
          list(method = "restyle", args = list("marker.opacity", op), label = paste0(op * 100, "%"))
        })
      )
    )
  ) %>%
  config(
    displayModeBar = TRUE,
    modeBarButtonsToAdd = c("lasso2d", "select2d")
  )

plotly3_SBP

```



**BKMR Results 2** {.tabset}
============
Row
-----------------------------------------------------------------------
### **Key Takeaway**
In the hemoglobin-A1C chart, levels dip as exposure to per- and polyfluoroalkyl substances (PFAS) increases, but the drop is not reliable enough to be sure it isn’t due to chance.

The glucose chart likewise shows no clear pattern with PFAS exposure and isn’t dependable enough to draw a firm conclusion.

Row
-----------------------------------------------------------------------
### **Hemoglobin-A1C (HA1C)**
```{r, echo=FALSE, warning=FALSE, message=FALSE}
bkmr_hemo_model <- readRDS(here::here("data", "bkmr_hemo_model.rds"))
risks.overall_hemo <- OverallRiskSummaries(fit = bkmr_hemo_model, qs = seq(0.25, 0.75, by = 0.05), q.fixed = 0.5)

bkmr_hemo_plot <- ggplot(data = risks.overall_hemo,  mapping = aes(
    x = quantile,
    y = est,
    ymin = est - 1.96 * sd,
    ymax = est + 1.96 * sd,
    text = paste0(
      "Mean: ", round(est, 2), "<br>",
      "95% CI: [", round(est - 1.96 * sd, 2), ", ", round(est + 1.96 * sd, 2), "]"
    )
  )
)  + geom_hline(yintercept = 0, lty = 2, col = "brown") + geom_pointrange() +    
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 8)) +
  xlab("Quantiles of Joint PFAS Exposures") +
  ylab("Overall Mixture Effect on Hemoglobin A1c")

plotly4_hemo <- ggplotly(bkmr_hemo_plot, tooltip = "text") %>% 
  layout(
    title = list(
      text = "Figure 4. Total Effect of PFAS Mixture on Hemoglobin A1C Levels",
      font = list(size = 12)
            ),
    xaxis = list(title = "Quantiles of Joint Exposure"),
    yaxis = list(title = "Overall Effect (No unit)"),
    updatemenus = list(
      list(
        type = "slider",
        active = 3,
        x = -0.1, xanchor = "left", y = 1.15, yanchor = "top",
        currentvalue = list(prefix = "Opacity: "),
        steps = lapply(seq(0.2, 1, by = 0.2), function(op) {
          list(method = "restyle", args = list("marker.opacity", op), label = paste0(op * 100, "%"))
        })
      )
    )
  ) %>%
  config(
    displayModeBar = TRUE,
    modeBarButtonsToAdd = c("lasso2d", "select2d")
  )

plotly4_hemo
```


### **Glucose**
```{r, echo=FALSE, warning=FALSE, message=FALSE}
bkmr_glucose_model <- readRDS(here::here("data", "bkmr_glucose_model.rds"))
risks.overall_glucose <- OverallRiskSummaries(fit = bkmr_glucose_model, qs = seq(0.25, 0.75, by = 0.05), q.fixed = 0.5)

bkmr_glucose_plot <- ggplot(data = risks.overall_glucose,  
    mapping = aes(
    x = quantile,
    y = est,
    ymin = est - 1.96 * sd,
    ymax = est + 1.96 * sd,
    text = paste0(
      "Mean: ", round(est, 2), "<br>",
      "95% CI: [", round(est - 1.96 * sd, 2), ", ", round(est + 1.96 * sd, 2), "]"
    )
  )
) + geom_hline(yintercept = 0, lty = 2, col = "brown") + geom_pointrange() +    
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 8)) +
  xlab("Quantiles of Joint PFAS Exposures") +
  ylab("Overall Mixture Effect on Glucose")

plotly5_glucose <- ggplotly(bkmr_glucose_plot, tooltip = "text") %>% 
  layout(
    title = list(
      text = "Figure 5. Total Effect of PFAS Mixture on Glucose Levels",
      font = list(size = 12)
            ),
    xaxis = list(title = "Quantiles of Joint Exposure"),
    yaxis = list(title = "Overall Effect (No unit)"),
    updatemenus = list(
      list(
        type = "slider",
        active = 3,
        x = -0.1, xanchor = "left", y = 1.15, yanchor = "top",
        currentvalue = list(prefix = "Opacity: "),
        steps = lapply(seq(0.2, 1, by = 0.2), function(op) {
          list(method = "restyle", args = list("marker.opacity", op), label = paste0(op * 100, "%"))
        })
      )
    )
  ) %>%
  config(
    displayModeBar = TRUE,
    modeBarButtonsToAdd = c("lasso2d", "select2d")
  )

plotly5_glucose
```


# **Dataset Description and Public Health Impact**

**Acknowledgement:** Although this is not a public dataset, I was given permission by Dr. Douglas Walker, Dr. Erin Ferranti, and Dr. Anne Dunlop to use the clinical and PFAS dataset provided that the participant information is deidentified.


**Dataset Decription:**
Sixty Black women from the Atlanta African American Maternal‑Child Cohort (Liang et al.) provided prenatal samples, and Dr. Erin Ferranti collected maternal venous blood 8–10 months postpartum along with sociodemographic, health, BMI, pregnancy complication/outcome, and postnatal clinical data. Serum samples were stored at –80°C and processed at Dr. Douglas Walker’s CLUES lab at Rollins School of Public Health. I used liquid chromatography–mass spectrometry (LC‑MS) to quantify 14 PFAS compounds in the sera. The raw PFAS concentrations were imported into R, cleaned, normalized with half‑minimum imputation, and processed. Finally, I merged the cleaned PFAS data with Dr. Ferranti’s clinical dataset to create the final analysis-ready database. 


**Public Health Impact:**
This project builds on previous research by focusing on the cardiometabolic risk factors in the Atlanta African American Maternal Child Cohort, an underrepresented population affected by maternal mortality and cardiometabolic risk. Ultimately, this project seeks to address critical research gaps in maternal and environmental health by highlighting and emphasizing the health challenges African American women face.
